---
title: "Experimentation Notebook in R"
author: "Umar Hassan"
date: "13/10/2020"
output: html_notebook
---

# My first notebook

Some basic stuff in R
```{r}
myname <- "Umar"
```

Let's store a number
```{r}
myage <- 24
```

## Importing data

To import data in R, you can use a function called read.csv().
```{r}
read.csv("Individual_claims_for_16_17.csv")
```

To store data that you can read in, you need to assign it to a variable like this:

```{r import1617}
#Import data from CSV
claims1617 <- read.csv("Individual_claims_for_16_17.csv")
claims1718 <- read.csv("Individual_claims_for_17_18.csv")
```

The variable we've created is a **dataframe** object.
```{r}
claims1718$Expense.Type
```

## Merge the datasets together for claims from 16/17 and 17/18.
```{r mergeclaims}
claims1618combined <- rbind(claims1617, claims1718)
```

##Export data

We can now export this as a new CSV as write.csv.

```{r}
write.csv(claims1618combined, "claims16to18.csv")
```

## Different types of variables

Touched on string and number variables, as well as df variables. 

Vectors are another type of variable and used to store a list.

This can be created using the c() function.

```{r}
listofages <- c(10,20,60,43)
listofnames <- c("Allen", "Paul", "Kevin", "Shawn")
listofbooleans <- c(TRUE, FALSE, TRUE, FALSE)
```

A dataframe is made up of a number of vectors (columns).
```{r}
mydataframe <- data.frame(listofnames, listofages, listofbooleans)
```

Use these functions to find out the properties of your vector.
```{r}
typeof(listofages)
length(listofnames)
is.vector(listofages)
class(listofbooleans)
```

Extract vector from dataframe using the $ sign.
```{r}
expensenames <- claims1617$MP.s.Name
```

```{r}
statuses <- claims1617$Status
```

```{r}
table(statuses)
```

I want to create a subset of my data and this is how you would do it.
```{r subset claims data from 1 April 2016 to 1 April 2017}
claims1Apr16 <- subset(claims1617, claims1617$Date == "01/04/2016")
claims31Mar17 <- subset(claims1617, claims1617$Date == "31/03/2017")
```

How do you create a basic barplot in R? This is how you would do it.
```{r basic barplot}
barplot(height = claims1618combined$Amount.Paid)
barplot(height = claims1618combined$Amount.Paid, ylab = "Expense amount paid", xlab = "Date")
```

A variable in R can be deleted like this.
```{r}
rm(claims_barplot)
```

You can delete as many variables as you like to make your environment workspace clean and tidy.

```{r}
rm(claims30Apr16, claims31Apr16)
```

