---
title: "visualise_expenses16_20"
author: "Umar Hassan"
date: "14 October 2020"
output: html_document
---

## Import csv data for individual claims for expenses covering 2016 to 2020

Import individual claims for expenses from 2018 to 2020.

```{r}
claims1819 <- read.csv("Individual_claims_for_18_19.csv")
claims1920 <- read.csv("Individual_claims_for_19_20.csv")
```

## Summary of expenses claimed for 1617

I want to know how many expenses were claimed by amount for 2018/19.

```{r}
summary(claims1819$Amount.Claimed)
```

## Summary of expenses claimed for 1920

I want to know how many expenses were claimed by amount for 2019/20.

```{r}
summary(claims1920$Amount.Claimed)
```

## Join the datasets together

I want to join the datasets for each expenses year from 18/19 to 19/20 using the rbind function.

```{r}
claimscombined_1820 <- rbind(claims1819, claims1920)
```

## Write the combined expenses file into a csv to access in Excel.

```{r}
write.csv(claimscombined_1820, "expenses2018to2020.csv")
```

## Create a subset, focusing on Liam Byrne MP (Lab, Hodge Hill)

```{r}
lbyrnemp_expenses <- subset(claimscombined_1820, claimscombined_1820$MP.s.Name == "Liam Byrne")
```

## Maximum expenses claimed for Liam Byrne MP (Lab, Hodge Hill)

How much did Liam Byrne MP claim in terms of maximum expenses?

```{r}
lbyrne_maxexp <- max(claimscombined_1820$Amount.Claimed)
lbyrne_maxexp
```

## Minimum expenses claimed for Liam Byrne MP (Lab, Hodge Hill)

```{r}
lbyrne_minexp <- min(claimscombined_1820$Amount.Claimed)
lbyrne_minexp
```

#Visualise the expense categories in ggplot2

Now, what we need to do is open up the tidyverse package, which includes ggplot2 to visualise our barplot for expenses.

```{r}
install.packages("tidyverse")#Only do this if you don't have the tidyverse package installed on your machine.
library(tidyverse)
library(scales)#This adds the currency to the data as we're visualising financial data in the form of MPs expenses.
```

Create the barplot

```{r}
lbyrne_barplot <- ggplot(lbyrnemp_expenses) + geom_bar(aes(x=Category), fill="#FFA500") + coord_flip() + labs(title = "How many times did Liam Byrne MP make an expense by category?", caption = "Source:Independent Parliamentary Standards Authority | Visualisation: Umar Hassan") + ylab("Category amount") + theme_set(theme_bw())
lbyrne_barplot
```