---
title: "westmids_drugsoutcomesaug20"
author: "Umar Hassan"
date: "15 October 2020"
output: html_document
---

##Import West Midlands Police crime data for August 2020 into RStudio.

Import west-mids-street data into R for August 2020.

```{r}
wmcrimesaug20 <- read.csv("2020-08-west-midlands-street.csv")
```

##Create simple pivot table in RStudio, focusing on Crime.type.

```{r - simple pivot table}
table(wmcrimesaug20$Crime.type)
```

## I want to drill in further and focus on a particular type of crime so I need to activate the tidyverse library to do this.

```{r}
install.packages("tidyverse")#Only do this if you don't have the tidyverse package installed on your machine.
library(tidyverse)
```

This is based upon a tutorial I found from [Glyn Mottershead's website]("https://www.glynmottershead.com/tutorial/tidyverse/pivot-tables-in-r/") on creating pivot tables using the tidyverse library to do it, using police data.

##Drill pivot table to focus on drugs

```{r}
drugs <- wmcrimesaug20 %>%
  filter(Crime.type == "Drugs") %>%
    rename("Outcome" = "Last.outcome.category", "Crimes" = "Crime.type") %>%
  mutate(Outcome = recode(Outcome,
    "Investigation complete; no suspect identified" = "No suspect identified")) %>%
  group_by(Outcome) %>%
  summarise(outcome_count = n()) %>%
  arrange(desc(outcome_count))
```

#Create csv file for drugs pivot table
```{r}
write.csv(drugs, "drugs_westmidspolaug20_pivot.csv")
```


## Visualise data using the ggplot2 package.

```{r}
drug_outcomes <-ggplot(drugs, aes(x= reorder(Outcome, outcome_count) ,outcome_count)) + geom_bar(stat = "identity", fill = "#0564FF") + labs(title = "Outcome for drugs crimes by category for West Midlands Police for August 2020", caption = "Source: West Midlands Police | Chart: Umar Hassan") + theme_set(theme_bw()) + xlab("") + ylab("") + coord_flip()
drug_outcomes
```

#Export image of barplot into png.image.

Create png.image for barplot for Drugs outcomes for West Mids police.

```{r}
ggsave("drug_outcomes.png", height= 6, width = 12, dpi=200)
```

